---
title: "Cooccurance of Columbidae (based on worldwide records) and resulting urban tolerance"
output: html_notebook
---

```{r setup}
library(dplyr)
library(tidyverse)
library(picante)
library(network)
```

Load data for our columbidae species present in all localities of the world
```{r}
all_occurance = read_csv(gzfile('../../data/bigquery/columbidae_birdlife__worldwide_ebird_records.csv.gz'))
```
```{r}
head(all_occurance)
```
```{r}
mapping = read_csv('../../data/bigquery/columbidae_birdlife_to_jetz__taxonomic_mapping.csv')
```

```{r}
all_occurance_jetz_pt1 = left_join(all_occurance, mapping, by = c("scientific_name" = "species_name"))
all_occurance_jetz = all_occurance_jetz_pt1 %>% group_by(locality_id, jetz_scientific_name) %>% summarise()
all_occurance_jetz
```

And then join to self using locality ID, which gives us all cooccuring species across the world, with urban tolerance.
```{r}
all_occurance_jetz2 = all_occurance_jetz
names(all_occurance_jetz2) = c('locality_id', 'jetz_scientific_name2')
cooccurance = right_join(all_occurance_jetz, all_occurance_jetz2)
cooccurance = cooccurance[cooccurance$jetz_scientific_name != cooccurance$jetz_scientific_name2,]
```

```{r}
head(cooccurance)
```

Remove locality so we end up with cooccuring species, along with number of patches recorded as cooccuring.
```{r}
cooccurance_details = cooccurance %>% group_by(jetz_scientific_name, jetz_scientific_name2) %>% summarise(recorded_patches = n())
cooccurance_details
```

Load in our regional pools
```{r}
regional_pools = read_csv('../../data/bigquery/columbidae_jetz__urban_locations.csv')
regional_pools = regional_pools[,c('jetz_scientific_name', 'city_id', 'present_in_city')]
regional_pools
```

Generate all cooccuring species in regional pools
```{r}
regional_pools2 = regional_pools
names(regional_pools2) = c('jetz_scientific_name2', 'city_id', 'present_in_city2')
regional_cooccurance = right_join(regional_pools, regional_pools2)
regional_cooccurance = regional_cooccurance[regional_cooccurance$jetz_scientific_name != regional_cooccurance$jetz_scientific_name2,]
head(regional_cooccurance)
```

```{r}
regional_cooccurance_pairs = regional_cooccurance %>% group_by(jetz_scientific_name, jetz_scientific_name2) %>% summarise(number_of_pools = n(), present_count = sum(present_in_city), present_count2 = sum(present_in_city2), both_present = sum(present_in_city & present_in_city2))
regional_cooccurance_pairs
```

Finally we can create all valid pairs of species that could be living in the same patches from our regional pools.
Here we also drop all species where neither pair is present in the city
```{r}
regional_patch_cooccurance = right_join(regional_cooccurance_pairs, cooccurance_details)
regional_patch_cooccurance = regional_patch_cooccurance[!is.na(regional_patch_cooccurance$number_of_pools),]
regional_patch_cooccurance = regional_patch_cooccurance[regional_patch_cooccurance$present_count >= regional_patch_cooccurance$present_count2,]
regional_patch_cooccurance = regional_patch_cooccurance[regional_patch_cooccurance$present_count > 0,]
regional_patch_cooccurance
```


In most of the records above, the number of times they are both present is the same that the number of times only the least dominant is presence.
The exceptions are:
```{r}
regional_patch_cooccurance[regional_patch_cooccurance$present_count2 != regional_patch_cooccurance$both_present,]
```

Problem records we have where they have the same tolerance (67 pairs of species):
```{r}
regional_patch_cooccurance[regional_patch_cooccurance$present_count == regional_patch_cooccurance$present_count2,]
```

Drop duplicate co-occuring pairs by creating a unique key from the two species names (should remove 67 rows)
```{r}
regional_patch_cooccurance = regional_patch_cooccurance %>% rowwise() %>%
  mutate(key = paste0(sort(c(jetz_scientific_name, jetz_scientific_name2)), collapse = '|'))

regional_patch_cooccurance = regional_patch_cooccurance[!duplicated(regional_patch_cooccurance$key),]

regional_patch_cooccurance
```

Work out MNTD between each species pair:
```{r}
tree <- read.tree("../../data/third_party/phylogeny_jetz__stage2_hackett_mcc_no_neg.tre")
```

```{r}
all_species = unique(c(regional_patch_cooccurance$jetz_scientific_name, regional_patch_cooccurance$jetz_scientific_name2))
tree_pruned_to_columbidae <- ladderize(drop.tip(tree, setdiff(tree$tip.label, all_species)))
```

```{r}
phydist <- cophenetic(tree_pruned_to_columbidae)
```


```{r}
regional_patch_cooccurance$mntd = NA

for(row_i in 1:nrow(regional_patch_cooccurance)) {
  community = data.frame(species1 = c(T, T), species2 = c(T, T))
  names(community) = list(as.character(regional_patch_cooccurance[[row_i,c('jetz_scientific_name')]]), as.character(regional_patch_cooccurance[[row_i,c('jetz_scientific_name2')]]))
  rownames(community) = c('1', '2')

  mntd.community.result <- ses.mntd(as.matrix(community), phydist, null.model="taxa.labels", abundance.weighted=FALSE, runs=99)
  
  regional_patch_cooccurance[row_i,]$mntd = mntd.community.result[1,]$mntd.obs
}

regional_patch_cooccurance
```

Calculate a status which indicates whether species are in the city together:
```{r}
regional_patch_cooccurance$cooccurance_status = 'Occasional'
regional_patch_cooccurance$cooccurance_status[regional_patch_cooccurance$present_count2 == 0] = 'Never'
regional_patch_cooccurance$cooccurance_status[regional_patch_cooccurance$present_count == regional_patch_cooccurance$present_count2] = 'Always'

regional_patch_cooccurance
```

```{r}
ggplot(regional_patch_cooccurance, aes(x = cooccurance_status, y = mntd)) + geom_boxplot()
```

```{r}
shapiro.test(regional_patch_cooccurance$mntd) 
shapiro.test(log(regional_patch_cooccurance$mntd))
shapiro.test(sqrt(regional_patch_cooccurance$mntd))
```

```{r}
nrow(regional_patch_cooccurance)
mean(regional_patch_cooccurance$mntd)
```

```{r}
regional_patch_cooccurance %>% group_by(cooccurance_status) %>% summarise(mean = mean(mntd), sd = sd(mntd), min = min(mntd), max = max(mntd), total = n())
```


```{r}
t.test(regional_patch_cooccurance$mntd[regional_patch_cooccurance$cooccurance_status == 'Never'], regional_patch_cooccurance$mntd[regional_patch_cooccurance$cooccurance_status == 'Always'])
```

```{r}
kruskal.test(cooccurance_status ~ mntd, data = regional_patch_cooccurance)
```

# Map Co-occurance using a network plot

## All co-occurance
```{r}
library(GGally)
```

```{r}
regional_patch_cooccurance
```

```{r}
regional_patch_cooccurance$number_of_pools_sqrt = as.integer(sqrt(regional_patch_cooccurance$number_of_pools))
regional_patch_cooccurance$cooccurance_ratio = regional_patch_cooccurance$both_present / regional_patch_cooccurance$number_of_pools

cooccurance_ratio_color_ramp <- colorRampPalette(c("red", "blue"))
cooccurance_ratio_palette = cooccurance_ratio_color_ramp(11)

get_ratio_colour = function(ratio) {
  cooccurance_ratio_palette[round(ratio * 10)]
}

color.gradient <- function(x, colors=c("red","yellow", "green"), colsteps=100) {
  return( colorRampPalette(colors) (colsteps) [ findInterval(x, seq(min(x),max(x), length.out=colsteps)) ] )
}

regional_patch_cooccurance$cooccurance_ratio_colour = color.gradient(regional_patch_cooccurance$cooccurance_ratio)
```


```{r}
ggnet2(regional_patch_cooccurance, label = T, edge.size = "number_of_pools_sqrt", edge.color = "cooccurance_ratio_colour")
ggsave("figures/worldwide_cooccurance.jpg", width=5112, height=5112, units = "px")
```

## Cooccurrance by realm

```{r}
head(regional_pools)
```

```{r}
cities_to_realms = read_csv('../../data/earth_engine/cities_to_realms.csv')
names(cities_to_realms) = c('city_id', 'realm')
cities_to_realms[order(cities_to_realms$realm),]
```
```{r}
cities_to_realms[cities_to_realms$city_id == 4355,]
```

```{r}
regional_pools_by_realm = left_join(cities_to_realms, regional_pools)
regional_pools_by_realm = regional_pools_by_realm[!is.na(regional_pools_by_realm$jetz_scientific_name),]
regional_pools_by_realm
```

```{r}
regional_pools_by_realm_summary = regional_pools_by_realm %>% group_by(realm, jetz_scientific_name) %>% summarise(species_presence_total_pools = n(), species_presence_total_urban = sum(present_in_city))
regional_pools_by_realm_summary
```

```{r}
regional_pools_by_realm2 = regional_pools_by_realm
names(regional_pools_by_realm2) = c('city_id', 'realm', 'jetz_scientific_name2', 'present_in_city2')
urban_cooccurance_by_realm = right_join(regional_pools_by_realm, regional_pools_by_realm2)
urban_cooccurance_by_realm = urban_cooccurance_by_realm[urban_cooccurance_by_realm$jetz_scientific_name != urban_cooccurance_by_realm$jetz_scientific_name2,]
head(urban_cooccurance_by_realm)
```

Restrict down to pairs of species that do occur
```{r}
possible_cooccurrance_by_realm = urban_cooccurance_by_realm %>% group_by(realm, jetz_scientific_name, jetz_scientific_name2) %>% summarise(cooccurance_total_pools = n(), cooccurance_both_urban = sum(present_in_city & present_in_city2))

cooccurrance_by_realm = right_join(cooccurance_details, possible_cooccurrance_by_realm)

cooccurrance_by_realm
```

Add additional columns for plot
```{r}
cooccurrance_by_realm$cooccurance_total_pools_sqrt = as.integer(sqrt(cooccurrance_by_realm$cooccurance_total_pools))
cooccurrance_by_realm$cooccurance_ratio = cooccurrance_by_realm$cooccurance_both_urban / cooccurrance_by_realm$cooccurance_total_pools
cooccurrance_by_realm$cooccurance_ratio_colour = color.gradient(cooccurrance_by_realm$cooccurance_ratio)

cooccurrance_by_realm
```

Summary data for each species for nodes
```{r}
species_presence_by_realm = regional_pools_by_realm %>% group_by(jetz_scientific_name, realm) %>% summarise(
  species_total_pools = n(), 
  species_total_cities = sum(present_in_city),
  species_urban_tolerance = sum(present_in_city) / n())

species_presence_by_realm$urban_tolerance_colour = color.gradient(species_presence_by_realm$species_urban_tolerance)
species_presence_by_realm
```

```{r}
unique(cooccurrance_by_realm$realm)
```

```{r}
cooccurrance_paleartic = cooccurrance_by_realm[cooccurrance_by_realm$realm == 'Palearctic',]
cooccurrance_paleartic
```

```{r}
cooccurrance_paleartic_net = network(cooccurrance_paleartic)
cooccurrance_paleartic_net
```

```{r}
species_paleartic = data.frame(species_presence_by_realm[species_presence_by_realm$realm == 'Palearctic',])
rownames(species_paleartic) = species_paleartic$jetz_scientific_name
species_paleartic
```

```{r}
ggnet2(cooccurrance_paleartic_net, label = T, edge.size = "cooccurance_total_pools_sqrt", edge.color = "cooccurance_ratio_colour", node.color = species_paleartic[,c('urban_tolerance_colour')], node.size = species_paleartic[,c('species_total_pools')])

ggsave("figures/paleartic_cooccurance.jpg", width=5112, height=5112, units = "px")
```





