---
title: "Phylogenetic signal for urban tolerance of Columbidae species"
output: html_notebook
---

```{r setup}
library(phytools)
library(dplyr)
library(tidyverse)
```

Load a maximum clade credibility tree from jetz
```{r}
tree <- read.tree("./phylogeny__stage2_hackett_mcc_no_neg.tre")
```

Load urban tolerance data
## Read in data exported from bigquery
```{yaml metadata all_columbidae_communities}
models:
  - name: all_columbidae_communities
    description: Regional and urban communities for all cities with a population of at least 250,000 from GHSL Data Package 2019 (Pesaresi et al. 2019), which contain at leasts 167 eBird checklists, and have at least one Columbidae species within their regional pools.
    columns:
      - name: city_id
        description: Vector ID from GHSL Data Package 2019 (Pesaresi et al. 2019)
      - name: city_name
        description: City name from Vector from GHSL Data Package 2019 (Pesaresi et al. 2019)
      - name: jetz_scientific_name
        description: The mapped name of each species whose distribution from BirdLife International’s species distribution maps (BirdLife International and Handbook of the Birds of the World 2020), overlaps with the city vector from GHSL Data Package 2019 (Pesaresi et al. 2019). Species mapped from eBird and Birdlife to JETZ taxonomy using birdlife alternative names and Avibase (Lepage 2011).
      - name: present_in_city
        description: True if the species is present on at least 5% of all valid checklists within the city
      - name: season_code 
        description: season code from BirdLife International and Handbook of the Birds of the World 2020
      - name: season
        description: Textual description of Birdlife's season code
      - name: origin_code 
        description: origin code from BirdLife International and Handbook of the Birds of the World 2020
      - name: origin
        description: Textual description of Birdlife's origin code
```

```{r read all_columbidae_presence}
all_columbidae_presence = read_csv('bigquery_export__columbidae_jetz_urban_locations.csv')
head(all_columbidae_presence)
```
Create urban tolerance for each species based on the number of cities / number of regional pools
```{r}
columbidae_urban_tolerance = all_columbidae_presence %>% group_by(jetz_scientific_name) %>% summarise(regional_pools = n(), urban_pools = sum(present_in_city), urban_tolerance = sum(present_in_city) / n())
head(columbidae_urban_tolerance)
```

Check for species from data not in tree
```{r}
columbidae_urban_tolerance[!columbidae_urban_tolerance$jetz_scientific_name %in% tree$tip.label,]
```

Number of species in analysis:
```{r}
nrow(columbidae_urban_tolerance)
```

Format data
```{r}
columbidae_input_species <- columbidae_urban_tolerance[columbidae_urban_tolerance$jetz_scientific_name %in% tree$tip.label,]
tree_pruned_to_columbidae <- ladderize(drop.tip(tree, setdiff(tree$tip.label, columbidae_urban_tolerance$jetz_scientific_name)))

columbidae_response <- columbidae_input_species$urban_tolerance
names(columbidae_response) <- columbidae_input_species$jetz_scientific_name

head(columbidae_response)
```

Is the a phylogenetic signal for urban tolerance in the columbidae family
```{r}
col_ut_phylo_signal <- phylosig(tree_pruned_to_columbidae, columbidae_response, method="lambda", test=TRUE)
col_ut_phylo_signal
```

Phylogenetic analysis containing all 160 species and produced a significant phylogenetic signal for urban tolerance (lambda: 0.16, LR(lambda=0): 2.37, p = 0.12)


Show the results for species
```{r}
col_continuous_trait_evolution <- contMap(tree_pruned_to_columbidae, columbidae_response)

colourCount <- length(col_continuous_trait_evolution$cols)
col_continuous_trait_evolution$cols[1:colourCount] <- rev(col_continuous_trait_evolution$col[1:colourCount])
```

```{r}
cte_all = plot(col_continuous_trait_evolution, outline=FALSE, fsize=0.3, lwd=1, leg.txt="Cities/Regional Pools", type = "fan")
cte_all
```

```{r}
jpeg('./figures/phylogentic_signal.jpg', width = 2056, height = 2056)
plot(col_continuous_trait_evolution, outline=FALSE, fsize=2, lwd=c(12, 8), leg.txt="Cities/Regional Pools", type = "fan")
dev.off()
```