---
title: "Phylogenetic signal for urban tolerance of Columbidae species"
output: html_notebook
---

```{r setup}
library(phytools)
library(dplyr)
library(tidyverse)
```

Load a maximum clade credibility tree from jetz
```{r}
tree <- read.tree("../../data/third_party/phylogeny_jetz__stage2_hackett_mcc_no_neg.tre")
```

## Read in data exported from bigquery
```{r read all_columbidae_presence}
all_columbidae_presence = read_csv('../../data/bigquery/columbidae_jetz__urban_locations.csv')
head(all_columbidae_presence)
```

Create urban tolerance for each species based on the number of cities / number of regional pools
```{r}
columbidae_urban_tolerance = all_columbidae_presence %>% group_by(jetz_scientific_name) %>% summarise(regional_pools = n(), urban_pools = sum(present_in_city), urban_tolerance = sum(present_in_city) / n())
head(columbidae_urban_tolerance)
```

Check for species from data not in tree
```{r}
columbidae_urban_tolerance[!columbidae_urban_tolerance$jetz_scientific_name %in% tree$tip.label,]
```

Number of species in analysis:
```{r}
nrow(columbidae_urban_tolerance)
```

Format data
```{r}
columbidae_input_species <- columbidae_urban_tolerance[columbidae_urban_tolerance$jetz_scientific_name %in% tree$tip.label,]
tree_pruned_to_columbidae <- ladderize(drop.tip(tree, setdiff(tree$tip.label, columbidae_urban_tolerance$jetz_scientific_name)))

columbidae_response <- columbidae_input_species$urban_tolerance
names(columbidae_response) <- columbidae_input_species$jetz_scientific_name

head(columbidae_response)
```

Is the a phylogenetic signal for urban tolerance in the columbidae family
```{r}
col_ut_phylo_signal <- phylosig(tree_pruned_to_columbidae, columbidae_response, method="lambda", test=TRUE)
col_ut_phylo_signal
```

Phylogenetic analysis containing all 160 species and produced a significant phylogenetic signal for urban tolerance (lambda: 0.16, LR(lambda=0): 2.37, p = 0.12)

Show the results for species
```{r}
col_continuous_trait_evolution <- contMap(tree_pruned_to_columbidae, columbidae_response)

colourCount <- length(col_continuous_trait_evolution$cols)
col_continuous_trait_evolution$cols[1:colourCount] <- rev(col_continuous_trait_evolution$col[1:colourCount])
```

```{r}
cte_all = plot(col_continuous_trait_evolution, outline=FALSE, fsize=0.3, lwd=1, leg.txt="Cities/Regional Pools", type = "fan")
cte_all
```

```{r}
jpeg('./figures/phylogentic_signal.jpg', width = 2056, height = 2056)
plot(col_continuous_trait_evolution, outline=FALSE, fsize=2, lwd=c(12, 8), leg.txt="Cities/Regional Pools", type = "fan")
dev.off()
```