---
title: "Urban Columbidae Communities"
output: html_notebook
---

```{r setup, message = F, warning = false}
library(dplyr)
library(purrr)
library(tidyverse)
library(tidyr)


library(phytools)
library(picante)
library(abdiv)
library(vegan)
library(fundiversity)
library(ggpubr)
library(gridExtra)
library(grid)
```

# Introduction
## Are urban columbidae communities strutured by habitat filtering or via competitive interactions?
- With competitive interactions we expect the phylogenetic and functional distance between species to higher than expected than within a random community.
- With habitat filtering we expect the phylogentic and functional distance between species to be lower than expected than within a random community.
(Swenson et al. 2007; Cavender-Bares et al. 2009).

## Recommended community metrics (Trisos et al. 2014)
- Functional diversity of all traits (Petchey and Gaston 2002)
- Convex Hull Volume (CHV) of all traits (a multivariate measure of the range of community values) (Cornwell et al. 2006)
- Mean taxon distance (MTD)  is mean pairwise phylogenetic distances between co-occurring species and is most sensitive to tree-wide patterns of phylogenetic clustering and evenness.
- Mean nearest taxon distance (MNTD) is mean pairwise phylogenetic distances separating each species from its closest co-occurring relative and is most sensitive to phylogenetic clustering or evenness at the tips of phylogeny.
- FD of each trait individually 
- Variance of each trait within community previously used to measure trait clustering
- SDNDr (standard deviation of distances between neighbouring species along a single trait axis, divided by the trait range of the community) is used to detect the regular spacing in species trait values predicted by competitive exclusion.

## Recommended traits (Trisos et al. 2014)
- Overall body size.
- Trophic - beak shape (beak length, width, depth) - measures competitive exclusion (e.g. over-dispersion) via accentuation of foraging differences
- Locomotory (wing, tail, tarsus length)

## Recommended null model (Milar et al. 2017)
Many metrics sensitive to species richness. Urban communities are governed by the surrounding regional pool. For each urban community, we can generate all of the possible communities with the same species richness as the urban community with species from the regional pool.  The average of all possible communities gives us a null model.

## Note
Expensive to create datasets are stored in cache files under ./cache.  To completely run the analysis from scratch ensure all files are removed from this directory first.

# Community data
We used all cities with a population of 250,000 or more, and that have at least 167 complete checklists that are more than 5 and less than 240 minutes in length and cover no more than 10km. We then included any Columbidae species in the urban community that appeared on at least 5% of all of these checklists recorded within the city.

This criteria comes from choosing all cities with at least 1000 checklists, and then testing 1000 times how many checklists are required to find all those species that appear on at least 5% of the total checklists. 269 cities were included in this test, and 167 was the maximum number of checklists required to find all min-5% species (which was in Auckland).

## Read in community data exported from bigquery
```{r read all_columbidae_communities}
all_columbidae_communities = read_csv('../../data/bigquery/columbidae_jetz__urban_locations.csv')
head(all_columbidae_communities)
```

# Site Selection
Our site metrics all compare species within communities, therefore we must have more than one species in each urban community for our analysis.
Our null model requires us to create a range of dissimilar communities with the same richness as the actual urban community using species from the regional pool, therefore we must have less than the number of species in our regional pool in our urban pool.

Here we exclude then two sets of urban communities that are interesting in their own right, (1) those with 1 or less species, and (2) those with all the species from the regional pool

```{r}
urban_communities_tmp = all_columbidae_communities %>% group_by(city_id, city_name) %>% summarise(regional_pool_size = n(), urban_pool_size = sum(present_in_city), number_of_regional_introduced_species = sum(origin == 'Introduced'), number_of_urban_introduced_species = sum(origin == 'Introduced' & present_in_city))
head(urban_communities_tmp)
```

## Cities that are excluded from remaining analysis
Communities with 1 or less pigeon species
```{r}
urban_communities_tmp[urban_communities_tmp$urban_pool_size < 2,]
```
Cities with no pigeons
```{r}
paste(urban_communities_tmp$city_name[urban_communities_tmp$urban_pool_size == 0], collapse = ', ')
```

Cities with a single pigeon but with more in regional pool
```{r}
paste(urban_communities_tmp$city_name[urban_communities_tmp$urban_pool_size == 1 & urban_communities_tmp$regional_pool_size > 1], collapse = ', ')
```
Cities with a single pigeon in city and regional pool
```{r}
paste(urban_communities_tmp$city_name[urban_communities_tmp$urban_pool_size == 1 & urban_communities_tmp$regional_pool_size == 1], collapse = ', ')
```

Cities with all species from regional pool where regional pool has more than 1 species
```{r}
urban_communities_tmp[urban_communities_tmp$urban_pool_size == urban_communities_tmp$regional_pool_size & urban_communities_tmp$regional_pool_size > 1,]
```

Cities that contain all regional species, when the number of regional species is greater than 1: 
Culiacán (6 species), Chihuahua(3), Rabat (4), Plymouth (4), Edinburgh (4), Stockholm (3), and Christchurch (2)

## Cities included in analysis
Selected sites:
```
models:
  - name: selected_urban_communities
    description: Sites that we will analyse that have more than 1 species, and less species than the entire regional pool
    columns:
      - name: city_id
        description: Vector ID from GHSL Data Package 2019 (Pesaresi et al. 2019)
      - name: city_name
        description: City name from Vector from GHSL Data Package 2019 (Pesaresi et al. 2019)
      - name: regional_pool_size
        description: the number of columbidae species in the city regional pool, e.g. number of species where the city vector overlaps the birdlife distribution vector
      - name: urban_pool_size
        description: the number of columbidae species that occur inside the city, e.g. the number of columbidae species that appear on at least 5% of all valid checklists within the city vector.
      - name: number_of_regional_introduced_species
        description: the number of columbidae species in the regional pool that are introduced, e.g. have origin 'Introduced' from the Birdlife distribution vector.
      - name: number_of_urban_introduced_species
        description: the number of columbidae species present inside the city that are introduced, e.g. have origin 'Introduced' from the Birdlife distribution vector.
```
```{r}
selected_urban_communities = urban_communities_tmp[urban_communities_tmp$urban_pool_size > 1 & urban_communities_tmp$urban_pool_size != urban_communities_tmp$regional_pool_size,]
nrow(selected_urban_communities)
head(selected_urban_communities)
```

And therefore our selected communities are:
```{r}
selected_columbidae_communities = all_columbidae_communities[all_columbidae_communities$city_id %in% selected_urban_communities$city_id,]
head(selected_columbidae_communities)
```

Total number of species across regional pools:
```{r}
length(unique(selected_columbidae_communities$jetz_scientific_name))
```

Total number of species across cities:
```{r}
length(unique(selected_columbidae_communities$jetz_scientific_name[selected_columbidae_communities$present_in_city]))
```

## Create a second set containing only Resident species
We have shown that non-resident species are less likely to be present in urban communities; however, this could be temporal sampling bias that means non-resident species are less likely to appear on at least 5% of all checklists. Therefore we should validate our findings against a second dataset that contains only resident species.

```{r}
all_resident_columbidae_communities = all_columbidae_communities[all_columbidae_communities$season == 'Resident',]
head(all_resident_columbidae_communities)
```

```{r}
resident_urban_communities_tmp = all_resident_columbidae_communities %>% group_by(city_id, city_name) %>% summarise(regional_pool_size = n(), urban_pool_size = sum(present_in_city))
head(resident_urban_communities_tmp)
```

Selected sites with resident only species:
```
models:
  - name: selected_resident_only_urban_communities
    description: Sites that we will analyse that have more than 1 species, and less species than the entire regional pool based on species being within their Residential range (and not just Breeding, Migration, or non-Breeding range).
    columns:
      - name: city_id
        description: Vector ID from GHSL Data Package 2019 (Pesaresi et al. 2019)
      - name: city_name
        description: City name from Vector from GHSL Data Package 2019 (Pesaresi et al. 2019)
      - name: regional_pool_size
        description: the number of columbidae species in the city regional pool, e.g. number of species where the city vector overlaps the birdlife distribution vector
      - name: urban_pool_size
        description: the number of columbidae species that occur inside the city, e.g. the number of columbidae species that appear on at least 5% of all valid checklists within the city vector.
```
```{r}
selected_resident_only_urban_communities = resident_urban_communities_tmp[resident_urban_communities_tmp$urban_pool_size > 1 & resident_urban_communities_tmp$urban_pool_size != resident_urban_communities_tmp$regional_pool_size,]
nrow(selected_resident_only_urban_communities)
head(selected_resident_only_urban_communities)
```

And therefore resident species at our selected urban communties are:
```{r}
selected_resident_columbidae_communities = all_resident_columbidae_communities[all_resident_columbidae_communities$city_id %in% selected_resident_only_urban_communities$city_id,]
head(selected_resident_only_urban_communities)
```

# Trait Data
Created in ./urban_traits/trait_axis.Rmd
```{r}
avonet_jetz = read_csv('../../data/output/urban_traits__trait_axis.csv')
```

# Calculate city wide community values

## First build out matrix of all urban communities
```{r}
species_in_cities = pivot_wider(selected_columbidae_communities[selected_columbidae_communities$present_in_city,c('jetz_scientific_name', 'present_in_city', 'city_id')], names_from = jetz_scientific_name, values_from = "present_in_city", values_fill = list(present_in_city = F))
species_in_cities = tibble::column_to_rownames(species_in_cities, var='city_id')
species_in_cities
```

and for resident only species
```{r}
resident_species_in_cities = pivot_wider(selected_resident_columbidae_communities[selected_resident_columbidae_communities$present_in_city,c('jetz_scientific_name', 'present_in_city', 'city_id')], names_from = jetz_scientific_name, values_from = "present_in_city", values_fill = list(present_in_city = F))
resident_species_in_cities = tibble::column_to_rownames(resident_species_in_cities, var='city_id')
resident_species_in_cities
```

## Read in phylogeny
Phylogenetic maximum clade tree based on a set of trees from Jetz et al. (2012) using the Hacket et al. (2008) backbone.
Trim tree to just species within our urban communities
```{r}
all_avian_species_tree <- read.tree("../../data/third_party/phylogeny_jetz__stage2_hackett_mcc_no_neg.tre")
columbidae_tree <- ladderize(drop.tip(all_avian_species_tree, setdiff(all_avian_species_tree$tip.label, selected_columbidae_communities$jetz_scientific_name)))
```

Check that we have all species in our tree - expect empty dataset
```{r}
selected_columbidae_communities[!selected_columbidae_communities$jetz_scientific_name %in% columbidae_tree$tip.label,]
```

## MNTD
Using picante package https://cran.r-project.org/web/packages/picante/index.html
Kembel, S.W., Cowan,P.D., Helmus, M.R., Cornwell, W.K., Morlon, H., Ackerly, D.D., Blomberg, S.P., Webb, C.O. (2010) Picante: {R} tools for integrating phylogenies and ecology. Bioinformatics, 26, 1463-1464
```{r get_mntd_for_communities}
get_mntd_for_communities = function(communities, phydist_urban_columbidae = cophenetic(columbidae_tree)) {
  urban_community_mntd_result = mntd(as.matrix(communities), phydist_urban_columbidae)
  data.frame(site = rownames(communities), mntd = urban_community_mntd_result)
}

get_mntd_for_communities(species_in_cities)
```

## MTD
Using picante package https://cran.r-project.org/web/packages/picante/index.html
Citation above.
```{r get_mtd_for_communities}
get_mtd_for_communities = function(communities, phydist_urban_columbidae = cophenetic(columbidae_tree)) {
  urban_community_mtd_result = mpd(as.matrix(communities), phydist_urban_columbidae)
  data.frame(site = rownames(communities), mtd = urban_community_mtd_result)
}

get_mtd_for_communities(species_in_cities)
```

## FD - Petchey and Gaston
From
https://github.com/NGSwenson/SwensonSESYNCWorkshop2017/blob/5a966564d6896ff0f41dc4ac0a6216f97b9d18e1/day.1.pm/tree.based.alpha.R

Petchey, O.L. and Gaston, K.J. (2002), Functional diversity (FD), species richness and community composition. Ecology Letters, 5: 402-411

Create dentrogram
```{r get_dentrogram_for_traits, fig.height=4}
analysis_all_traits = c('mass', 'locomotory_trait', 'trophic_trait')

get_traits_data_frame = function(trait_names) {
  trait_names = prepend('jetz_scientific_name', trait_names)
  traits = avonet_jetz[,trait_names]
  traits = tibble::column_to_rownames(traits, var='jetz_scientific_name')
  traits
}

get_dentrogram_for_traits = function(required_traits_for_dentro) {
  traits = get_traits_data_frame(required_traits_for_dentro)
  hclust(dist(traits,method="euclidean"), method="average")
}

dentro = get_dentrogram_for_traits(analysis_all_traits)
plot(dentro)
```

Petchey and Gaston's FD
```{r get_fd_for_traits}
get_fd_for_traits = function(required_traits_for_pd, communities, dentorgram = get_dentrogram_for_traits(required_traits_for_pd)) {
  result = pd(communities, as.phylo(dentorgram))
  data.frame(site = rownames(result), fd = result$PD)
}

urban_community_fd_result = get_fd_for_traits(analysis_all_traits, species_in_cities)
head(urban_community_fd_result)
```

```{r}
urban_community_fd_mass_result = get_fd_for_traits(c('locomotory_trait'), species_in_cities)
head(urban_community_fd_mass_result)
```

## Trait Variance
### FRic - Convex Hull of all traits
Using fundiversity package https://cran.r-project.org/web/packages/fundiversity/index.html
Grenié M, Gruson H (2022). fundiversity: Easy Computation of Functional Diversity Indices. doi:10.5281/zenodo.4761754, R package version 1.1.1

```{r get_convex_hull}
get_convex_hull = function(trait_names, communities) {
  analysis_trait_values = get_traits_data_frame(trait_names)
  fd_fric(analysis_trait_values, communities)
}

get_convex_hull(analysis_all_traits)
```

### Single Trait Variance
How to get names of species in urban community:
```{r}
species_in_514 = species_in_cities[c('514'),]
species_names_in_514 = names(species_in_514)[which(species_in_514 == 1, arr.ind=T)[, "col"]]
species_names_in_514
```

How to get values of traits for species in urban community, and then var of those traits:
```{r trait_values}
trait_values = function(trait, community_row) {
  species_names_in_community = names(community_row)[which(community_row == 1, arr.ind=T)[, "col"]]
  traits_for_species_in_community = avonet_jetz[avonet_jetz$jetz_scientific_name %in% species_names_in_community, ]
  traits_for_species_in_community[[c(trait)]]
}

trait_values('mass', species_in_cities[c('514'),])
var(trait_values('mass', species_in_cities[c('514'),]))
```

Turn into a function:
```{r single_trait_variance}
single_trait_variance = function(trait, communities) {
  variance_of_trait = data.frame()
  for(row in rownames(communities)) {
    community_row = communities[c(row),]
    species_names_in_community = names(community_row)[which(community_row == 1, arr.ind=T)[, "col"]]
    
    result = data.frame(site = row, variance = var(trait_values(trait, community_row)))
    variance_of_trait = rbind(variance_of_trait, result)
  }
  
  variance_of_trait
}

single_trait_variance('mass', species_in_cities)
```

## SDNDr
Standard deviation of distances between neighbouring species along a single trait axis, divided by the trait range of the community.

How to calculate SDNDr using trait values for an urban community:
```{r}
print('trait values:')
sorted_loco_traits_in_1291 = sort(trait_values('locomotory_trait', species_in_cities[c('1291'),]))
sorted_loco_traits_in_1291 

print('diff in neighbouring trait values:')
diff(sorted_loco_traits_in_1291)

print('SDNDr:')
sd(diff(sorted_loco_traits_in_1291)) / (max(sorted_loco_traits_in_1291) - min(sorted_loco_traits_in_1291))
```

Turn into a function:
```{r SDNDr}
SDNDr = function(trait, communities) {
  sdndr_of_trait = data.frame()
  for(row in rownames(communities)) {
    community_row = communities[c(row),]
    
    sorted_trait_values = sort(trait_values(trait, community_row))
    
    sdndr_value = sd(diff(sorted_trait_values)) / (max(sorted_trait_values) - min(sorted_trait_values))
    
    sdndr_of_trait = rbind(sdndr_of_trait, data.frame(site = row, sdndr = sdndr_value))
  }
  
  sdndr_of_trait
}

SDNDr('locomotory_trait', species_in_cities)
```

## Put it all together
```{r get_community_metrics}
get_community_metrics = function(communities, included_traits = analysis_all_traits) {

  mntd_result = get_mntd_for_communities(communities)
  mtd_result = get_mtd_for_communities(communities)
  fd_all_traits = get_fd_for_traits(included_traits, communities)
  convex_hull = get_convex_hull(included_traits, communities)
  
  result = right_join(right_join(mntd_result, mtd_result), right_join(fd_all_traits, convex_hull))
  
  for (trait in included_traits) {
    col_name = function(value) {
      paste(trait, value, sep = '_')
    }
  
    fd_single_trait = get_fd_for_traits(trait, communities)
    var_single_trait = single_trait_variance(trait, communities)
    sdndr_single_trait = SDNDr(trait, communities)
    
    single_trait_result = right_join(right_join(fd_single_trait, var_single_trait), sdndr_single_trait)
    names(single_trait_result) = c('site', col_name('fd'), col_name('var'), col_name('sdndr'))
    result = right_join(result, single_trait_result)
  }
  
  result
}
```

```{r}
urban_community_values = get_community_metrics(species_in_cities)
urban_community_values
```

```{r}
urban_resident_only_community_values = get_community_metrics(resident_species_in_cities)
urban_resident_only_community_values
```

# Calculate the null models
Our null model is every possible community of species from the regional pool with the same richness as the urban pool

Find all species in regional pool:
```{r}
test_city_id = 1786 # Manchester

all_columbidae_communities$jetz_scientific_name[all_columbidae_communities$city_id == test_city_id]
```

```{r}
all_columbidae_communities$jetz_scientific_name[all_columbidae_communities$city_id == test_city_id]
```

Find urban richness
```{r}
selected_urban_communities$urban_pool_size[selected_urban_communities$city_id == test_city_id]
```

Function to build every possible community
```{r}
create_null_model_communities = function(city_id, urban_richness, regional_species) {
  regional_richness = length(regional_species)
  
  null_model_sites = expand.grid(rep(list(TRUE:FALSE), regional_richness))
  null_model_sites_species_richness =  data.frame(species_richness = pmap_dbl(null_model_sites, sum))
  null_model_sites = cbind(null_model_sites, null_model_sites_species_richness)
  
  null_model_sites = null_model_sites[null_model_sites$species_richness == urban_richness,1:regional_richness]
  colnames(null_model_sites) = regional_species
  null_model_sites
}

create_null_model_communities(
  test_city_id, 
  selected_urban_communities$urban_pool_size[selected_urban_communities$city_id == test_city_id], 
  all_columbidae_communities$jetz_scientific_name[all_columbidae_communities$city_id == test_city_id]
)
```

What were the results for the actual urban community in Manchester?
```{r}
urban_community_values[urban_community_values$site == test_city_id,]
```

Test creating community metrics for our null model
```{r}
test_null_models = create_null_model_communities(
  test_city_id, 
  selected_urban_communities$urban_pool_size[selected_urban_communities$city_id == test_city_id],
  all_columbidae_communities$jetz_scientific_name[all_columbidae_communities$city_id == test_city_id]
)

test_null_model_community_values = get_community_metrics(test_null_models)
test_null_model_community_values
```

```{r}
test_city_id2 = 8
get_community_metrics(
  create_null_model_communities(
    test_city_id2, 
    selected_urban_communities$urban_pool_size[selected_urban_communities$city_id == test_city_id2],
    all_columbidae_communities$jetz_scientific_name[all_columbidae_communities$city_id == test_city_id2]
  )
)
```

# Compare Urban Communities to regional null models
```{r normalise}
normalise = function(actual_value, min_value, max_value) {
  (actual_value - min_value) / (max_value - min_value)
}

normalise(1.5, 1, 2)
```

```{r standardise}
standardise = function(actual_value, mean_value, sd) {
  (actual_value - mean_value) / sd
}

standardise(1.5, 1, 0.25)
```

Test creating normalised and standardised FRic
```{r}
property = 'FRic'
actual_value = urban_community_values[urban_community_values$site == test_city_id,c(property)]
null_model_values = test_null_model_community_values[,c(property)]

print(paste('Normalised', property, 'at site', test_city_id))
normalise(actual_value, min(null_model_values), max(null_model_values))
print(paste('Standardised', property, 'at site', test_city_id))
standardise(actual_value, mean(null_model_values), sd(null_model_values))
```

Create function to test actual value against all possible null model values
```{r}
compare_community_to_null_model = function(metric_name, actual_metric_value, null_model_metric_values) {
  min_value = min(null_model_metric_values)
  max_value = max(null_model_metric_values)
  mean_value = mean(null_model_metric_values)
  sd_value = sd(null_model_metric_values)
  
  result = data.frame(
    normalised = normalise(actual_metric_value, min_value, max_value),
    stanardised = standardise(actual_metric_value, mean_value, sd_value),
    actual = actual_metric_value,
    min = min_value,
    max = max_value,
    mean = mean_value,
    sd = sd_value
  )
  
  col_name = function(value) {
      paste(metric_name, value, sep = '_')
  }
  
  colnames(result) = c(col_name('normalised'), col_name('standardised'), col_name('actual'), col_name('min'), col_name('max'), col_name('mean'), col_name('sd'))
  result
}

compare_community_to_null_model(property, actual_value, null_model_values)
```

The metrics we want to test:
```{r}
all_metrics_to_test = colnames(urban_community_values)[-1]
all_metrics_to_test
```

Compare the actual values to the null model values:
```{r}
create_null_model_test_values = function(city_id, urban_richness, regional_species, actual_values) {
  null_models = create_null_model_communities(city_id, urban_richness, regional_species)
  null_model_community_values = get_community_metrics(null_models)
  
  result = data.frame(city_id = city_id)
  
  for (metric in all_metrics_to_test) {
    result = cbind(result, 
                   compare_community_to_null_model(metric, actual_values[,c(metric)], null_model_community_values[,c(metric)]))
  }
  result
}

create_null_model_test_values(
  test_city_id, 
  selected_urban_communities$urban_pool_size[selected_urban_communities$city_id == test_city_id],
  all_columbidae_communities$jetz_scientific_name[all_columbidae_communities$city_id == test_city_id], 
  urban_community_values[urban_community_values$site == test_city_id,]
)
```


```{r}
create_null_model_test_values(
  test_city_id2, 
  selected_urban_communities$urban_pool_size[selected_urban_communities$city_id == test_city_id2],
  all_columbidae_communities$jetz_scientific_name[all_columbidae_communities$city_id == test_city_id2],
  urban_community_values[urban_community_values$site == test_city_id2,]
)
```

And for resident only species
```{r}
create_null_model_test_values(
  test_city_id2, 
  selected_resident_only_urban_communities$urban_pool_size[selected_resident_only_urban_communities$city_id == test_city_id2],
  all_resident_columbidae_communities$jetz_scientific_name[all_resident_columbidae_communities$city_id == test_city_id2],
  urban_resident_only_community_values[urban_resident_only_community_values$site == test_city_id2,]
)
```


```{r}
ALL_SPECIES_RESULT__CACHE_FILE = './cache/community_assembly_all_species_result.csv'

if(!file.exists(ALL_SPECIES_RESULT__CACHE_FILE)) {
  urban_all_species_community_result = data.frame()
  
  for (loop_all_city_id in selected_urban_communities$city_id) {
    city_null_models_all_species = create_null_model_test_values(
      loop_all_city_id, 
      selected_urban_communities$urban_pool_size[selected_urban_communities$city_id == loop_all_city_id],
      all_columbidae_communities$jetz_scientific_name[all_columbidae_communities$city_id == loop_all_city_id],
      urban_community_values[urban_community_values$site == loop_all_city_id,]
    )
    
    urban_all_species_community_result = rbind(urban_all_species_community_result, city_null_models_all_species)
  }
  
  urban_all_species_community_result = right_join(urban_all_species_community_result, selected_urban_communities)
  
  write_csv(urban_all_species_community_result, ALL_SPECIES_RESULT__CACHE_FILE)
} else {
  urban_all_species_community_result = read_csv(ALL_SPECIES_RESULT__CACHE_FILE)
}

head(urban_all_species_community_result)
```

```{r}
RESIDENT_SPECIES_RESULT__CACHE_FILE = './cache/community_assembly_resident_species_result.csv'
urban_resident_only_species_community_result = NULL

if (!file.exists(RESIDENT_SPECIES_RESULT__CACHE_FILE)) {
  urban_resident_only_species_community_result = data.frame()
  
  for (loop_resident_city_id in selected_resident_only_urban_communities$city_id) {
    city_null_models_resident_species = create_null_model_test_values(
      loop_resident_city_id, 
      selected_resident_only_urban_communities$urban_pool_size[selected_resident_only_urban_communities$city_id == loop_resident_city_id],
      all_resident_columbidae_communities$jetz_scientific_name[all_resident_columbidae_communities$city_id == loop_resident_city_id],
      urban_resident_only_community_values[urban_resident_only_community_values$site == loop_resident_city_id,]
    )
    
    urban_resident_only_species_community_result = rbind(urban_resident_only_species_community_result, city_null_models_resident_species)
  }
  
  urban_resident_only_species_community_result = right_join(urban_resident_only_species_community_result, selected_urban_communities)
  
  write_csv(urban_resident_only_species_community_result, RESIDENT_SPECIES_RESULT__CACHE_FILE)
} else {
  urban_resident_only_species_community_result = read_csv(RESIDENT_SPECIES_RESULT__CACHE_FILE)
}

head(urban_resident_only_species_community_result)
```

# Analysis
Total number of cities in analysis
```{r}
nrow(urban_all_species_community_result)
```

Total number of cities in resident only species analysis
```{r}
nrow(urban_resident_only_species_community_result)
```

## Analysis of all traits across entire community

### MNTD
Pairwise phylogenetic distances separating each species from its closest co-occurring relative and is most sensitive to phylogenetic clustering or evenness at the tips of phylogeny.

```{r}
fmt_p_value = function(test_result) {
  ifelse(test_result$p.value > 0.001, paste('=', round(test_result$p.value, 3)), '< 0.001')
}

geom_standardised_histogram = function(name, gg, test_result, mean) {
    gg + geom_histogram(color = "#000000", fill = "#0099F8", binwidth = 0.25) +
    geom_vline(aes(xintercept = 0), color = "#000000", size = 1.25) +
    geom_vline(aes(xintercept = -1), color = "#000000", size = 1, linetype = "dashed") +
    geom_vline(aes(xintercept = 1), color = "#000000", size = 1, linetype = "dashed") + 
    ylab("Number of cities") + xlab("Standardised Response") + ylim(c(0, 150)) + xlim(c(-3, 3)) +
    labs(title = name, subtitle=paste('(V=', test_result$statistic, ', p-value ', fmt_p_value(test_result), ', mean = ', round(mean, 2), ')', sep=''))
}

geom_normalised_histogram = function(name, gg, test_result, mean) {
  gg + geom_histogram(color = "#000000", fill = "#0099F8", binwidth = 0.1) +
  geom_vline(aes(xintercept = 0.5), color = "#000000", size = 1.25) +
  geom_vline(aes(xintercept = 0), color = "#000000", size = 1, linetype = "dashed") +
  geom_vline(aes(xintercept = 1), color = "#000000", size = 1, linetype = "dashed") + 
  ylab("Number of cities") + xlab("Normalised Response") + ylim(c(0, 200)) +
  labs(title = name, subtitle = paste('(V=', test_result$statistic, ', p-value ', fmt_p_value(test_result), ', mean = ', round(mean, 2), ')', sep=''))
}
```

#### All species
Standardised 
```{r}
std_mntd_analysis_plot = geom_standardised_histogram(
  'MNTD', 
  ggplot(urban_all_species_community_result, aes(mntd_standardised)), 
  wilcox.test(urban_all_species_community_result$mntd_standardised, mu = 0), 
  mean(urban_all_species_community_result$mntd_standardised)
)
std_mntd_analysis_plot
```

Normalised
```{r}
norm_mntd_analysis_plot = geom_normalised_histogram(
  'MNTD', 
  ggplot(urban_all_species_community_result, aes(mntd_normalised)), 
  wilcox.test(urban_all_species_community_result$mntd_normalised, mu = 0.5), 
  mean(urban_all_species_community_result$mntd_normalised)
)
norm_mntd_analysis_plot
```

#### Resident only species
Normalised
```{r}
norm_mntd_analysis_plot_ro = geom_normalised_histogram(
  'MNTD', 
  ggplot(urban_resident_only_species_community_result, aes(mntd_normalised)), 
  wilcox.test(urban_resident_only_species_community_result$mntd_normalised, mu = 0.5), 
  mean(urban_resident_only_species_community_result$mntd_normalised)
)
norm_mntd_analysis_plot_ro
```

### MTD
Mean pairwise phylogenetic distances between co-occurring species and is most sensitive to tree-wide patterns of phylogenetic clustering and evenness.

#### All species
Standardised 
```{r}
std_mtd_analysis_plot = geom_standardised_histogram(
  'MTD', 
  ggplot(urban_all_species_community_result, aes(mtd_standardised)), 
  wilcox.test(urban_all_species_community_result$mtd_standardised, mu = 0), 
  mean(urban_all_species_community_result$mtd_standardised)
)
std_mtd_analysis_plot
```

Normalised
```{r}
norm_mtd_analysis_plot = geom_normalised_histogram(
  'MTD', 
  ggplot(urban_all_species_community_result, aes(mtd_normalised)), 
  wilcox.test(urban_all_species_community_result$mtd_normalised, mu = 0.5), 
  mean(urban_all_species_community_result$mtd_normalised)
)
norm_mtd_analysis_plot
```


#### Resident only species
Normalised
```{r}
norm_mtd_analysis_plot_ro = geom_normalised_histogram(
  'MTD', 
  ggplot(urban_resident_only_species_community_result, aes(mtd_normalised)), 
  wilcox.test(urban_resident_only_species_community_result$mtd_normalised, mu = 0.5), 
  mean(urban_resident_only_species_community_result$mtd_normalised)
)
norm_mtd_analysis_plot_ro
```

### All Traits FD

#### All species
Standardised 
```{r}
std_fd_analysis_plot = geom_standardised_histogram(
  'FD', 
  ggplot(urban_all_species_community_result, aes(fd_standardised)), 
  wilcox.test(urban_all_species_community_result$fd_standardised, mu = 0), 
  mean(urban_all_species_community_result$fd_standardised)
)
std_fd_analysis_plot
```

Normalised:
```{r}
norm_fd_analysis_plot = geom_normalised_histogram(
  'FD', 
  ggplot(urban_all_species_community_result, aes(fd_normalised)), 
  wilcox.test(urban_all_species_community_result$fd_normalised, mu = 0.5), 
  mean(urban_all_species_community_result$fd_normalised)
)
norm_fd_analysis_plot
```

#### Resident only species
Normalised
```{r}
norm_fd_analysis_plot_ro = geom_normalised_histogram(
  'FD', 
  ggplot(urban_resident_only_species_community_result, aes(fd_normalised)), 
  wilcox.test(urban_resident_only_species_community_result$fd_normalised, mu = 0.5), 
  mean(urban_resident_only_species_community_result$fd_normalised)
)
norm_fd_analysis_plot_ro
```


### FRic
Number of cities that we could calculate FRic for, need 4 or more species (as require more species than traits)

#### All species
Standardised 
```{r}
std_fric_analysis_plot = geom_standardised_histogram(
  'FRic', 
  ggplot(urban_all_species_community_result, aes(FRic_standardised)), 
  wilcox.test(urban_all_species_community_result$FRic_standardised, mu = 0), 
  mean(urban_all_species_community_result$FRic_standardised, na.rm = T)
)
std_fric_analysis_plot
```

Normalised
```{r}
norm_fric_analysis_plot = geom_normalised_histogram(
  'FRic', 
  ggplot(urban_all_species_community_result, aes(FRic_normalised)), 
  wilcox.test(urban_all_species_community_result$FRic_normalised, mu = 0.5), 
  mean(urban_all_species_community_result$FRic_normalised, na.rm = T)
)
norm_fric_analysis_plot
```

#### Resident only species
Normalised
```{r}
norm_fric_analysis_plot_ro = geom_normalised_histogram(
  'FRic', 
  ggplot(urban_resident_only_species_community_result, aes(FRic_normalised)), 
  wilcox.test(urban_resident_only_species_community_result$FRic_normalised, mu = 0.5), 
  mean(urban_resident_only_species_community_result$FRic_normalised, na.rm = T)
)
norm_fric_analysis_plot_ro
```

## Single Trait Metrics

Number of cities we can calculate SDNDr for - require at least 3 species
```{r}
nrow(urban_all_species_community_result[!is.na(urban_all_species_community_result$mass_sdndr_actual),])
```

### Locomotory Trait

#### FD

##### All species
Standardised 
```{r}
std_loco_trait_fd_analysis_plot = geom_standardised_histogram(
  'Locomotory Trait - FD', 
  ggplot(urban_all_species_community_result, aes(locomotory_trait_fd_standardised)), 
  wilcox.test(urban_all_species_community_result$locomotory_trait_fd_standardised, mu = 0), 
  mean(urban_all_species_community_result$locomotory_trait_fd_standardised, na.rm = T)
)
std_loco_trait_fd_analysis_plot
```

Normalised
```{r}
norm_loco_trait_fd_analysis_plot = geom_normalised_histogram(
  'Locomotory Trait - FD', 
  ggplot(urban_all_species_community_result, aes(locomotory_trait_fd_normalised)), 
  wilcox.test(urban_all_species_community_result$locomotory_trait_fd_normalised, mu = 0.5), 
  mean(urban_all_species_community_result$locomotory_trait_fd_normalised, na.rm = T)
)
norm_loco_trait_fd_analysis_plot
```

##### Resident only species
Normalised
```{r}
norm_loco_trait_fd_analysis_plot_ro = geom_normalised_histogram(
  'Locomotory Trait - FD', 
  ggplot(urban_resident_only_species_community_result, aes(locomotory_trait_fd_normalised)), 
  wilcox.test(urban_resident_only_species_community_result$locomotory_trait_fd_normalised, mu = 0.5), 
  mean(urban_resident_only_species_community_result$locomotory_trait_fd_normalised, na.rm = T)
)
norm_loco_trait_fd_analysis_plot_ro
```

#### VAR

##### All species
Standardised 
```{r}
std_loco_trait_var_analysis_plot = geom_standardised_histogram(
  'Locomotory Trait - VAR', 
  ggplot(urban_all_species_community_result, aes(locomotory_trait_var_standardised)), 
  wilcox.test(urban_all_species_community_result$locomotory_trait_var_standardised, mu = 0), 
  mean(urban_all_species_community_result$locomotory_trait_var_standardised, na.rm = T)
)
std_loco_trait_var_analysis_plot
```

Normalised
```{r}
norm_loco_trait_var_analysis_plot = geom_normalised_histogram(
  'Locomotory Trait - VAR', 
  ggplot(urban_all_species_community_result, aes(locomotory_trait_var_normalised)), 
  wilcox.test(urban_all_species_community_result$locomotory_trait_var_normalised, mu = 0.5), 
  mean(urban_all_species_community_result$locomotory_trait_var_normalised, na.rm = T)
)
norm_loco_trait_var_analysis_plot
```

##### Resident only species
Normalised
```{r}
norm_loco_trait_var_analysis_plot_ro = geom_normalised_histogram(
  'Locomotory Trait - VAR', 
  ggplot(urban_resident_only_species_community_result, aes(locomotory_trait_var_normalised)), 
  wilcox.test(urban_resident_only_species_community_result$locomotory_trait_var_normalised, mu = 0.5), 
  mean(urban_resident_only_species_community_result$locomotory_trait_var_normalised, na.rm = T)
)
norm_loco_trait_var_analysis_plot_ro
```

#### SDNDr

##### All species
Standardised 
```{r}
std_loco_trait_sdndr_analysis_plot = geom_standardised_histogram(
  'Locomotory Trait - SDNDr', 
  ggplot(urban_all_species_community_result, aes(locomotory_trait_sdndr_standardised)), 
  wilcox.test(urban_all_species_community_result$locomotory_trait_sdndr_standardised, mu = 0), 
  mean(urban_all_species_community_result$locomotory_trait_sdndr_standardised, na.rm = T)
)
std_loco_trait_sdndr_analysis_plot
```

Normalised
```{r}
norm_loco_trait_sdndr_analysis_plot = geom_normalised_histogram(
  'Locomotory Trait - SDNDr', 
  ggplot(urban_all_species_community_result, aes(locomotory_trait_sdndr_normalised)), 
  wilcox.test(urban_all_species_community_result$locomotory_trait_sdndr_normalised, mu = 0.5), 
  mean(urban_all_species_community_result$locomotory_trait_sdndr_normalised, na.rm = T)
)
norm_loco_trait_sdndr_analysis_plot
```

##### Resident only species
Normalised
```{r}
norm_loco_trait_sdndr_analysis_plot_ro = geom_normalised_histogram(
  'Locomotory Trait - SDNDr', 
  ggplot(urban_resident_only_species_community_result, aes(locomotory_trait_sdndr_normalised)), 
  wilcox.test(urban_resident_only_species_community_result$locomotory_trait_sdndr_normalised, mu = 0.5), 
  mean(urban_resident_only_species_community_result$locomotory_trait_sdndr_normalised, na.rm = T)
)
norm_loco_trait_sdndr_analysis_plot_ro
```

### Trophic Trait

#### FD

##### All species
Standardised 
```{r}
std_trophic_trait_fd_analysis_plot = geom_standardised_histogram(
  'Trophic Trait - FD', 
  ggplot(urban_all_species_community_result, aes(trophic_trait_fd_standardised)), 
  wilcox.test(urban_all_species_community_result$trophic_trait_fd_standardised, mu = 0), 
  mean(urban_all_species_community_result$trophic_trait_fd_standardised, na.rm = T)
)
std_trophic_trait_fd_analysis_plot
```

Normalised
```{r}
norm_trophic_trait_fd_analysis_plot = geom_normalised_histogram(
  'Trophic Trait - FD', 
  ggplot(urban_all_species_community_result, aes(trophic_trait_fd_normalised)), 
  wilcox.test(urban_all_species_community_result$trophic_trait_fd_normalised, mu = 0.5), 
  mean(urban_all_species_community_result$trophic_trait_fd_normalised, na.rm = T)
)
norm_trophic_trait_fd_analysis_plot
```

##### Resident only species
Normalised
```{r}
norm_trophic_trait_fd_analysis_plot_ro = geom_normalised_histogram(
  'Trophic Trait - FD', 
  ggplot(urban_resident_only_species_community_result, aes(trophic_trait_fd_normalised)), 
  wilcox.test(urban_resident_only_species_community_result$trophic_trait_fd_normalised, mu = 0.5), 
  mean(urban_resident_only_species_community_result$trophic_trait_fd_normalised, na.rm = T)
)
norm_trophic_trait_fd_analysis_plot_ro
```

#### VAR

##### All species
Standardised 
```{r}
std_trophic_trait_var_analysis_plot = geom_standardised_histogram(
  'Trophic Trait - VAR', 
  ggplot(urban_all_species_community_result, aes(trophic_trait_var_standardised)), 
  wilcox.test(urban_all_species_community_result$trophic_trait_var_standardised, mu = 0), 
  mean(urban_all_species_community_result$trophic_trait_var_standardised, na.rm = T)
)
std_trophic_trait_var_analysis_plot
```

Normalised
```{r}
norm_trophic_trait_var_analysis_plot = geom_normalised_histogram(
  'Trophic Trait - VAR', 
  ggplot(urban_all_species_community_result, aes(trophic_trait_var_normalised)), 
  wilcox.test(urban_all_species_community_result$trophic_trait_var_normalised, mu = 0.5), 
  mean(urban_all_species_community_result$trophic_trait_var_normalised, na.rm = T)
)
norm_trophic_trait_var_analysis_plot
```

##### Resident only species
Normalised
```{r}
norm_trophic_trait_var_analysis_plot_ro = geom_normalised_histogram(
  'Trophic Trait - VAR', 
  ggplot(urban_resident_only_species_community_result, aes(trophic_trait_var_normalised)), 
  wilcox.test(urban_resident_only_species_community_result$trophic_trait_var_normalised, mu = 0.5), 
  mean(urban_resident_only_species_community_result$trophic_trait_var_normalised, na.rm = T)
)
norm_trophic_trait_var_analysis_plot_ro
```

#### SDNDr

##### All species
Standardised 
```{r}
std_trophic_trait_sdndr_analysis_plot = geom_standardised_histogram(
  'Trophic Trait - SDNDr', 
  ggplot(urban_all_species_community_result, aes(trophic_trait_sdndr_standardised)), 
  wilcox.test(urban_all_species_community_result$trophic_trait_sdndr_standardised, mu = 0), 
  mean(urban_all_species_community_result$trophic_trait_sdndr_standardised, na.rm = T)
)
std_trophic_trait_sdndr_analysis_plot
```

Normalised
```{r}
norm_trophic_trait_sdndr_analysis_plot = geom_normalised_histogram(
  'Trophic Trait - SDNDr', 
  ggplot(urban_all_species_community_result, aes(trophic_trait_sdndr_normalised)), 
  wilcox.test(urban_all_species_community_result$trophic_trait_sdndr_normalised, mu = 0.5), 
  mean(urban_all_species_community_result$trophic_trait_sdndr_normalised, na.rm = T)
)
norm_trophic_trait_sdndr_analysis_plot
```

##### Resident only species
Normalised
```{r}
norm_trophic_trait_sdndr_analysis_plot_ro = geom_normalised_histogram(
  'Trophic Trait - SDNDr', 
  ggplot(urban_resident_only_species_community_result, aes(trophic_trait_sdndr_normalised)), 
  wilcox.test(urban_resident_only_species_community_result$trophic_trait_sdndr_normalised, mu = 0.5), 
  mean(urban_resident_only_species_community_result$trophic_trait_sdndr_normalised, na.rm = T)
)
norm_trophic_trait_sdndr_analysis_plot_ro
```

### Mass

#### FD

##### All species
Standardised 
```{r}
std_mass_fd_analysis_plot = geom_standardised_histogram(
  'Mass - FD', 
  ggplot(urban_all_species_community_result, aes(mass_fd_standardised)), 
  wilcox.test(urban_all_species_community_result$mass_fd_standardised, mu = 0), 
  mean(urban_all_species_community_result$mass_fd_standardised, na.rm = T)
)
std_mass_fd_analysis_plot
```

Normalised
```{r}
norm_mass_fd_analysis_plot = geom_normalised_histogram(
  'Mass - FD', 
  ggplot(urban_all_species_community_result, aes(mass_fd_normalised)), 
  wilcox.test(urban_all_species_community_result$mass_fd_normalised, mu = 0.5), 
  mean(urban_all_species_community_result$mass_fd_normalised, na.rm = T)
)
norm_mass_fd_analysis_plot
```

##### Resident only species
Normalised
```{r}
norm_mass_fd_analysis_plot_ro = geom_normalised_histogram(
  'Mass - FD', 
  ggplot(urban_resident_only_species_community_result, aes(mass_fd_normalised)), 
  wilcox.test(urban_resident_only_species_community_result$mass_fd_normalised, mu = 0.5), 
  mean(urban_resident_only_species_community_result$mass_fd_normalised, na.rm = T)
)
norm_mass_fd_analysis_plot_ro
```

#### VAR

##### All species
Standardised 
```{r}
std_mass_var_analysis_plot = geom_standardised_histogram(
  'Mass - VAR', 
  ggplot(urban_all_species_community_result, aes(mass_var_standardised)), 
  wilcox.test(urban_all_species_community_result$mass_var_standardised, mu = 0), 
  mean(urban_all_species_community_result$mass_var_standardised, na.rm = T)
)
std_mass_var_analysis_plot
```

Normalised
```{r}
norm_mass_var_analysis_plot = geom_normalised_histogram(
  'Mass - VAR', 
  ggplot(urban_all_species_community_result, aes(mass_var_normalised)), 
  wilcox.test(urban_all_species_community_result$mass_var_normalised, mu = 0.5), 
  mean(urban_all_species_community_result$mass_var_normalised, na.rm = T)
)
norm_mass_var_analysis_plot
```

##### Resident only species
Normalised
```{r}
norm_mass_var_analysis_plot_ro = geom_normalised_histogram(
  'Mass - VAR', 
  ggplot(urban_resident_only_species_community_result, aes(mass_var_normalised)), 
  wilcox.test(urban_resident_only_species_community_result$mass_var_normalised, mu = 0.5), 
  mean(urban_resident_only_species_community_result$mass_var_normalised, na.rm = T)
)
norm_mass_var_analysis_plot_ro
```

#### SDNDr

##### All species
Standardised 
```{r}
std_mass_sdndr_analysis_plot = geom_standardised_histogram(
  'Mass - SDNDr', 
  ggplot(urban_all_species_community_result, aes(mass_sdndr_standardised)), 
  wilcox.test(urban_all_species_community_result$mass_sdndr_standardised, mu = 0), 
  mean(urban_all_species_community_result$mass_sdndr_standardised, na.rm = T)
)
std_mass_sdndr_analysis_plot
```

Normalised
```{r}
norm_mass_sdndr_analysis_plot = geom_normalised_histogram(
  'Mass - SDNDr', 
  ggplot(urban_all_species_community_result, aes(mass_sdndr_normalised)), 
  wilcox.test(urban_all_species_community_result$mass_sdndr_normalised, mu = 0.5), 
  mean(urban_all_species_community_result$mass_sdndr_normalised, na.rm = T)
)
norm_mass_sdndr_analysis_plot
```

##### Resident only species
Normalised
```{r}
norm_mass_sdndr_analysis_plot_ro = geom_normalised_histogram(
  'Mass - SDNDr', 
  ggplot(urban_resident_only_species_community_result, aes(mass_sdndr_normalised)), 
  wilcox.test(urban_resident_only_species_community_result$mass_sdndr_normalised, mu = 0.5), 
  mean(urban_resident_only_species_community_result$mass_sdndr_normalised, na.rm = T)
)
norm_mass_sdndr_analysis_plot_ro
```

## Create Figures

### All Traits

Standardised
```{r, fig.height=3, fig.width=3}
ggarrange(std_mntd_analysis_plot, std_mtd_analysis_plot, std_fric_analysis_plot, std_fd_analysis_plot, ncol = 2, nrow = 2)
```

Normalised
```{r, fig.height=3, fig.width=3}
ggarrange(norm_mntd_analysis_plot, norm_mtd_analysis_plot, norm_fric_analysis_plot, norm_fd_analysis_plot, ncol = 2, nrow = 2)
```

Strong evidence for competitive exclusion shaping urban communities, three of the four metrics are significantly higher than would be expected by random.  The only metric that is not higher than expected nor significant is FRic. Could this be because FRic only includes urban communities with 3 or more species?

We can check the normalised response for FD against urban community size to see if there is any evidence of larger communities being less likely to be shaped by competitive exclusion. This shows that larger pools do have a much lower mean for normalised FD response and therefore the FRic result could be down to only larger pools being included.  More species could well be present when more habitat is available in the city, and thus there is sufficient habitat to help species avoid getting involved in competitive interactions.

```{r}
ggplot(urban_all_species_community_result, aes(y = fd_normalised, x = as.factor(urban_pool_size))) + geom_boxplot()
```
```{r, fig.height=3, fig.width=3}
blank <- grid.rect(gp=gpar(col="white"))

point_plot = function(points_aes) {
  ggplot(urban_all_species_community_result, points_aes) + 
    geom_point(size = 0.5, aes(color = as.factor(urban_pool_size))) + 
    scale_x_continuous(position = 'top') +
    guides(color=guide_legend(title="Urban richness", nrow=1)) +
    theme_bw()
}

ggarrange(
  point_plot(aes(x = fd_normalised, y = mntd_normalised)) + ylab("MNTD") + xlab("FD"),
  point_plot(aes(x = FRic_normalised, y = mntd_normalised)) + rremove("ylab") + xlab("FRic"),
  point_plot(aes(x = mtd_normalised, y = mntd_normalised)) + rremove("ylab") + xlab("MTD"),
  point_plot(aes(x = fd_normalised, y = mtd_normalised)) + ylab("MTD") + rremove("xlab"),
  point_plot(aes(x = FRic_normalised, y = mtd_normalised)) + rremove("ylab") + rremove("xlab"),
  blank,
  point_plot(aes(x = fd_normalised, y = FRic_normalised)) + ylab("FRic") + rremove("xlab"),
  blank,
  blank
  , ncol = 3, nrow = 3, common.legend = T, legend = "bottom")
```

```{r}
jpeg('./figures/community_assembly_community.jpg', width = 1200, height = 800)
ggarrange(norm_mntd_analysis_plot + theme_bw(), norm_mtd_analysis_plot + theme_bw(), norm_fric_analysis_plot + theme_bw(), norm_fd_analysis_plot + theme_bw(), ncol = 2, nrow = 2)
dev.off()
```

```{r}
jpeg('./figures/community_assembly_community_compared.jpg', width = 1200, height = 800)
ggarrange(
  point_plot(aes(x = fd_normalised, y = mntd_normalised)) + ylab("MNTD") + xlab("FD"),
  point_plot(aes(x = FRic_normalised, y = mntd_normalised)) + rremove("ylab") + xlab("FRic"),
  point_plot(aes(x = mtd_normalised, y = mntd_normalised)) + rremove("ylab") + xlab("MTD"),
  point_plot(aes(x = fd_normalised, y = mtd_normalised)) + ylab("MTD") + rremove("xlab"),
  point_plot(aes(x = FRic_normalised, y = mtd_normalised)) + rremove("ylab") + rremove("xlab"),
  blank,
  point_plot(aes(x = fd_normalised, y = FRic_normalised)) + ylab("FRic") + rremove("xlab"),
  blank,
  blank
  , ncol = 3, nrow = 3, common.legend = T, legend = "bottom")
dev.off()
```

Normalised - resident only species
```{r, fig.height=3, fig.width=3}
ggarrange(norm_mntd_analysis_plot_ro, norm_mtd_analysis_plot_ro, norm_fric_analysis_plot_ro, norm_fd_analysis_plot_ro, ncol = 2, nrow = 2)
```

### Single Traits
Standardised
```{r, fig.height=6, fig.width=6}
ggarrange(
  std_loco_trait_fd_analysis_plot, std_loco_trait_var_analysis_plot, std_loco_trait_sdndr_analysis_plot,
  std_trophic_trait_fd_analysis_plot, std_trophic_trait_var_analysis_plot, std_trophic_trait_sdndr_analysis_plot,
  std_mass_fd_analysis_plot, std_mass_var_analysis_plot, std_mass_sdndr_analysis_plot,
  ncol = 3, nrow = 3)
```

Normalised
```{r, fig.height=6, fig.width=6}
ggarrange(
  norm_loco_trait_fd_analysis_plot, norm_loco_trait_var_analysis_plot, norm_loco_trait_sdndr_analysis_plot,
  norm_trophic_trait_fd_analysis_plot, norm_trophic_trait_var_analysis_plot, norm_trophic_trait_sdndr_analysis_plot,
  norm_mass_fd_analysis_plot, norm_mass_var_analysis_plot, norm_mass_sdndr_analysis_plot,
  ncol = 3, nrow = 3)
```

Higher than expected values for the locomotory trait FD and mass FD, along with lower than expected values for the locomotory trait SDNDr and mass SDNDr give some evidence that both these traits are selected under competitive exclusion.  VAR in both cases gives the opposite view, this is the single trait version of FRic.
While the trophic trait has a higher than expected FD, its SDNDr value is fairly close to the mean and therefore this trait may not be selected under competitive exclusion.
Also, Trisos et al. (2014) discourage the use of multi-pattern metrics such as variance and range, which are sensitive to both clustering and over-dispersion, in single-pattern, one-tailed hypothesis tests.

```{r}
jpeg('./figures/community_assembly_single_traits.jpg', width = 1200, height = 800)
ggarrange(
  norm_loco_trait_fd_analysis_plot + theme_bw(), norm_loco_trait_var_analysis_plot + theme_bw(), norm_loco_trait_sdndr_analysis_plot + theme_bw(),
  norm_trophic_trait_fd_analysis_plot + theme_bw(), norm_trophic_trait_var_analysis_plot + theme_bw(), norm_trophic_trait_sdndr_analysis_plot + theme_bw(),
  norm_mass_fd_analysis_plot + theme_bw(), norm_mass_var_analysis_plot + theme_bw(), norm_mass_sdndr_analysis_plot + theme_bw(),
  ncol = 3, nrow = 3)
dev.off()
```

Locomotory metrics compared
```{r}
ggarrange(
  point_plot(aes(x = locomotory_trait_var_normalised, y = locomotory_trait_fd_normalised)) + ylab("FD") + xlab("VAR"),
  point_plot(aes(x = locomotory_trait_sdndr_normalised, y = locomotory_trait_fd_normalised)) + rremove("ylab") + xlab("SDNDr"),
  ncol = 2, nrow = 1, common.legend = T, legend = "bottom")
```

Trophic metrics compared
```{r}
ggarrange(
  point_plot(aes(x = trophic_trait_var_normalised, y = trophic_trait_fd_normalised)) + ylab("FD") + xlab("VAR"),
  point_plot(aes(x = trophic_trait_sdndr_normalised, y = trophic_trait_fd_normalised)) + rremove("ylab") + xlab("SDNDr"),
  ncol = 2, nrow = 1, common.legend = T, legend = "bottom")
```

Mass metrics compared
```{r}
ggarrange(
  point_plot(aes(x = mass_var_normalised, y = mass_fd_normalised)) + ylab("FD") + xlab("VAR"),
  point_plot(aes(x = mass_sdndr_normalised, y = mass_fd_normalised)) + rremove("ylab") + xlab("SDNDr"),
  ncol = 2, nrow = 1, common.legend = T, legend = "bottom")
```

Normalised - resident only species
```{r, fig.height=6, fig.width=6}
ggarrange(
  norm_loco_trait_fd_analysis_plot_ro, norm_loco_trait_var_analysis_plot_ro, norm_loco_trait_sdndr_analysis_plot_ro,
  norm_trophic_trait_fd_analysis_plot_ro, norm_trophic_trait_var_analysis_plot_ro, norm_trophic_trait_sdndr_analysis_plot_ro,
  norm_mass_fd_analysis_plot_ro, norm_mass_var_analysis_plot_ro, norm_mass_sdndr_analysis_plot_ro,
  ncol = 3, nrow = 3)
```

# Impact of Introduced species

## Is there any difference in MNTD or FD of cities with introduced species?
```{r}
wilcox.test(urban_all_species_community_result$fd_normalised ~ as.factor(urban_all_species_community_result$number_of_urban_introduced_species))
```

```{r}
ggplot(urban_all_species_community_result, aes(y = fd_normalised, x = as.factor(number_of_urban_introduced_species))) + geom_boxplot()
```

```{r}
wilcox.test(urban_all_species_community_result$mntd_normalised ~ as.factor(urban_all_species_community_result$number_of_urban_introduced_species))
```

```{r}
ggplot(urban_all_species_community_result, aes(y = mntd_normalised, x = as.factor(number_of_urban_introduced_species))) + geom_boxplot()
```

# Investigate by Realm

```{r}
head(urban_all_species_community_result)
```

Load in realms
```{r}
cities_to_realms = read_csv('../../data/earth_engine/cities_to_realms.csv')
names(cities_to_realms) = c('city_id', 'realm')
head(cities_to_realms[order(cities_to_realms$realm),])
```
```{r}
by_realm_urban_all_species_community_result = full_join(cities_to_realms, urban_all_species_community_result)
by_realm_urban_all_species_community_result$urban_richness = as.factor(by_realm_urban_all_species_community_result$urban_pool_size)
```

## MNTD

```{r, fig.height=6, fig.width=6}
ggplot(by_realm_urban_all_species_community_result, aes(x = mntd_normalised, fill = urban_richness)) + 
  geom_histogram(position='stack', color = "#000000", binwidth = 0.1) +
  geom_vline(aes(xintercept = 0.5), color = "#000000", size = 1.25) +
  geom_vline(aes(xintercept = 0), color = "#000000", size = 1, linetype = "dashed") +
  geom_vline(aes(xintercept = 1), color = "#000000", size = 1, linetype = "dashed") + 
  ylab("Number of cities") + xlab("Normalised Response") + 
  ylim(c(0, 50)) +
  labs(title = "MNTD") + 
  facet_wrap( ~ realm) + 
  theme(legend.position="bottom")
```

```{r, fig.height=6, fig.width=6}
ggplot(by_realm_urban_all_species_community_result, aes(x = mntd_normalised, fill = as.factor(number_of_urban_introduced_species))) + 
  geom_histogram(position='stack', color = "#000000", binwidth = 0.1) +
  geom_vline(aes(xintercept = 0.5), color = "#000000", size = 1.25) +
  geom_vline(aes(xintercept = 0), color = "#000000", size = 1, linetype = "dashed") +
  geom_vline(aes(xintercept = 1), color = "#000000", size = 1, linetype = "dashed") + 
  ylab("Number of cities") + xlab("Normalised Response") + 
  ylim(c(0, 50)) +
  labs(title = "MNTD") + 
  facet_wrap( ~ realm) + 
  theme(legend.position="bottom")
```

## FD

```{r, fig.height=6, fig.width=6}
ggplot(by_realm_urban_all_species_community_result, aes(x = fd_normalised, fill = urban_richness)) + 
  geom_histogram(color = "#000000", position='stack', binwidth = 0.1) +
  geom_vline(aes(xintercept = 0.5), color = "#000000", size = 1.25) +
  geom_vline(aes(xintercept = 0), color = "#000000", size = 1, linetype = "dashed") +
  geom_vline(aes(xintercept = 1), color = "#000000", size = 1, linetype = "dashed") + 
  ylab("Number of cities") + xlab("Normalised Response") + 
  ylim(c(0, 50)) +
  labs(title = "FD") + 
  facet_wrap( ~ realm) + 
  theme(legend.position="bottom")
```
## FRic

```{r, fig.height=6, fig.width=6}
ggplot(by_realm_urban_all_species_community_result, aes(x = FRic_normalised, fill = urban_richness)) + 
  geom_histogram(color = "#000000", position='stack', binwidth = 0.1) +
  geom_vline(aes(xintercept = 0.5), color = "#000000", size = 1.25) +
  geom_vline(aes(xintercept = 0), color = "#000000", size = 1, linetype = "dashed") +
  geom_vline(aes(xintercept = 1), color = "#000000", size = 1, linetype = "dashed") + 
  ylab("Number of cities") + xlab("Normalised Response") + 
  ylim(c(0, 50)) +
  labs(title = "FRic") + 
  facet_wrap( ~ realm) + 
  theme(legend.position="bottom")
```

